Artificial ecosystem selection reveals relationships between microbiome composition and ecosystem function

### 2023-06-23 manuscript/manuscript.Rmd
remove header
update with edits for reviewers

Microbiomes mediate important ecosystem functions, yet it has proven difficult
to determine the relationship between microbiome composition and the rate of
ecosystem functions. This challenge remains because it is difficult to
manipulate microbiome composition directly, we often cannot know _a priori_ which
microbiome members influence the rate of an ecosystem function, and
microbiomes can covary strongly with other drivers of ecosystem function, such
as the environment. We propose that artificial ecosystem selection can overcome
these challenges. Artificial ecosystem selection involves creating replicate
ecosystems, selecting replicates with a desired ecosystem-level trait, and
using the microbiomes from the selected ecosystems to inoculate a new set of
ecosystems. We performed artificial ecosystem selection on replicate soil
microbiomes to select for increased soil CH~4~ oxidation rate. We observed a
strong response to selection with a `r myround(per_slope, 1)`\% increase in
CH~4~ oxidation rate per passage. We
estimated that `r myround(tidy(rh_model)[[2, 2]]*100, 1)`\% of the variation in
CH~4~ oxidation
rate in these soils can be attributed to microbiome variation. We also found
that selection did not enrich for known CH~4~ oxidizers; instead, `r length(da_fams)` families not
known to oxidize CH~4~, including *Fimbriimonadaceae*, *Cytophagaceae*, and
*Diplorickettsiaceae*, were enriched by selection. This result is in contrast to the typical assumption that
the rate of an ecosystem function is limited by the final step in the associated
microbial pathway. Our study demonstrates that variation in microbiome
composition can contribute to variation in the rate of ecosystem function and
that manipulating microbiome composition could be a viable strategy for managing
ecosystem functions.

@@

In this study, we sought to build on the observations of comparative and
manipulative studies by applying a different approach to the question of whether
microbiome variation contributes to variation in the rate of an ecosystem
function. We used artificial ecosystem selection to select for microbiomes that
performed a greater rate of ecosystem function [@swenson2000; @blouin2015;
@panke-buisse2015]. We then tested whether variation in the microbiome
contributed to variation in the rate of ecosystem function and identified
microbiome attributes that might explain this relationship. There are several
potential advantages to this approach for documenting the direct contribution of
the microbiome to variation in ecosystem function and for investigating the
mechanisms underlying those relationships. By repeated passaging of microbiomes
in a common environment, we can weaken the covariance between microbes and the
environment by repeatedly diluting the influence of variation in abiotic
conditions. In addition, our approach eliminates the need to generate microbiome
variation through methods that are confounded with biomass or cell size. Lastly,
by comparing our artificially selected community to a control community
resulting from random selection, we can both control for changes in the
environment over time and identify genes or taxa that are associated with the
ecosystem function under selection.

@@

We performed an artificial ecosystem selection experiment 
(*sensu* @swenson2000) by passaging replicate soil microbiomes. The trait we
selected on was CH~4~ oxidation rate. This experiment had a selection line and a
control line with twelve jars each for a total of twenty-four jars per passage.
The selection line underwent positive selection where the two or three jars
with the highest CH~4~ oxidation rate were homogenized to inoculate the next set
of jars. The control line underwent random selection where an equal number
of jars as the selection line were chosen at random to inoculate the next set of
jars. The number of jars chosen was based on the distribution of fluxes among
the positive jars, i.e., we chose the top three jars unless only two jars had
considerably greater CH~4~ oxidation rates based on visual inspection of
histograms. The experiment was carried out over five passages until a
significant divergence in functional rates was observed between the two
selection treatments based on a difference of slopes.

The initial soil microbiome was sampled from the top 10 cm of an upland mineral
soil under a deciduous forest ecosystem near the University of Oregon campus in
Eugene, OR, USA. Incubations were performed in 500 mL mason jars with rubber
septa installed in the lids. Each jar was sterilized with 70\% ethanol to which
was added 45 g of autoclaved potting mix, 5 g of living soil, and 3.5 mL of
sterile deionized water to bring them to 60\% of field capacity. The potting
mix was Lane Potting Mix from Lane Forest Products (Eugene, OR, USA). Each jar
was then capped and injected with 4.3 mL of 99\% CH~4~ to bring
the headspace concentration to approximately 1000 ppm CH~4~. To create the two
treatment groups, twenty-four jars were created in an identical manner and then
randomly assigned to either the selection or control line. Jars were
flushed and respiked with CH~4~ twice per week to maintain aerobic conditions
and elevated CH~4~ concentrations and were incubated at ambient temperature for
approximately four weeks. Methane oxidation rates were determined at the end of
the incubation period and selection was performed. For each treatment, the
selected jars were homogenized and 5 g of the homogenized soil was used to
inoculate the next set of jars, which represents a 10% subsampling. The next set
of jars were created in an identical manner to the first generation with fresh
autoclaved potting mix and the same moisture and CH~4~ content.

@@

Methane oxidation rates were determined after flushing and spiking jars to
approximately 1000 ppm CH~4~. Headspace samples of 1 mL were collected from
each jar immediately after spiking and then at time points 3, 6, 24, and 48
hours for a 5-point curve. Samples were immediately injected into a SRI model
8610C gas chromatograph equipped with a flame ionization detector (SRI
Instruments, Torrance, CA, USA) to determine the headspace CH~4~ concentration.
We applied a first-order exponential decay function to determine the rate
constant (k, units = d^−1^; i.e., dCH~4~/dt = k[CH~4~]) of the exponential
decrease in CH~4~ Oxidation rates are presented as the additive inverse of
$k$ (i.e., $-k$) so that a more positive value represents a greater
oxidation rate. The jars selected in passage two for the positive selection treatment had
the lowest CH~4~ oxidation rate of the twelve jars due to a calculation error
in the rate constant. Despite this error, we still observed a strong
response to selection demonstrating the power of this technique. In addition,
this likely reintroduced diversity lost during the selection process,
potentially increasing the variation available to be selected upon, as
is recommended for microbiome selection experiments [@sanchez2021].

@@

Bioinformatics processing was performed in 'R' [@rcoreteam2018].  Demultiplexed
sequencing reads were denoised using 'DADA2' to generate a table of amplicon
sequence variants (ASVs) [@callahan2016]. Taxonomic assignment was performed
using the Ribosomal Database Project naive Bayesian classifier [@wang2007]. The
presence of contaminants was evaluated using both the prevalence and frequency
methods from 'DECONTAM' [@davis2018]. Decontam identified 
`r length(which(contamdf$contaminant))` potential contaminants based on
prevalence and frequency. Visual inspection of abundance-concentration plots
indicated that `r length(contaminants)` of these were likely contaminants and
these ASVs were removed. Amplicon sequence variants that were assigned
chloroplast or mitochondria taxonomy were removed prior to analysis.

@@

Statistical analyses were performed in 'R' [@rcoreteam2018]. To determine
whether there was a significant change in CH~4~ oxidation rate as a response to
selection, we tested a difference in slopes between the selection and control
lines. Residuals did not meet the assumptions of constant variance
and normal distribution. Therefore, CH~4~ oxidation rates were log~10~
transformed prior to analysis. First, we tested if there was a difference of
slopes between the selection line and the control based on the interaction
between passage and treatment. To test the interaction, we fit two nested models
with and without the interaction term and compared them using an F-test with the
'anova' function. We then present the slopes for each treatment, which
represented the change in CH~4~ oxidation rate per passage as a response to
selection.

We estimated the proportion of variation in CH~4~ oxidation rate due to
variation in the microbiome as the regression of divergence between the positive
line and the control on the cumulative selection differential [@falconer1996].
Divergence was calculated as the mean CH~4~ oxidation rate of the positive
treatment minus the mean CH~4~ oxidation rate of the control in each
passage. The selection differential was calculated as the difference between
the mean of the three selected jars and the mean of all twelve jars in a passage.
Cumulative selection differential was calculated as the sum of the selection
differential from all preceding selection events. We then regressed cumulative
divergence on cumulative selection differential using the 'lm' function.

@@

To identify taxa that responded to selection on CH~4~ oxidation rate, we tested
differential abundance between the two treatments in passage 5. We first
grouped ASVs at the family level. Any ASVs that lacked a family-level
taxonomic assignment were grouped at a higher taxonomic level. We then subset the
samples in Passage 5 and removed all families with a prevalence of less than
10\% in either treatment. We used three methods for testing differential
abundance: ANCOM-II, ALDEx2, and CORNCOB 
[@kaul2017; @lin2020; @fernandes2014; @martin2020]. We then identified the
consensus taxa that were significant with all three tests and plotted their
relative abundances. For ANCOM-II, we used the 'ancom' function in the
'ANCOM-BC' package with a cutoff of W = 0.7 [@kaul2017; @lin2020]. For ALDEx2,
we used the 'aldex' function in the 'ALDEx2' package with Welch's t-test and we
used an effect size of 1 as our significance threshold [@fernandes2014].
Finally, we used CORNCOB with the 'differentialTest' function in the 'corncob'
package with the Wald test and without bootstrapping [@martin2020]. Lastly, to
test differentially abundant methanotrophs, we subset all ASVs within
methanotrophic families and tested their differential abundance aggregated at
the family and genus level using 'corncob'.

@@

To estimate the proportion of variation in CH~4~ oxidation rate due to
variation in microbiome composition, we regressed divergence between the
positive selection treatment and the control against the cumulative selection
differential (`r figures("herit", display = "cite")`). This estimate is analogous
to estimates of "microbiability" from the animal breeding literature, which
quantifies the
variation in a host trait that is due to microbiome variation [@difford2016]. 
The slope of the regression of divergence on cumulative selection differential
provides an estimate of realized microbiability (h^2^ &pm; SE), which in our
experiment was `r myround(tidy(rh_model)[[2, 2]], 2)` &pm;
`r myround(tidy(rh_model)[[2, 3]], 2)`, though this was not significant (F~`r glance(rh_model)$df`,`r glance(rh_model)$df.residual`~ = 
`r myround(glance(rh_model)$statistic, 2)`, 
p = `r myround(glance(rh_model)$p.value, 2)`).

@@

Median ASV richness decreased from `r myround(rare_median[1, 2], 1)` (`r myround(rare_median[1, 3], 1)`) in passage 2 to `r myround(rare_median[2, 2], 1)` (`r myround(rare_median[2, 3], 1)`) in passage 5 (Kruskal-Wallace test: $\chi^2$ = 
`r myround(kt[1,1], 1)`, df = `r kt[1,3]`, p < 0.001; pairwise Wilcoxon test: p < 0.001). However, there was no
difference in richness between the selection treatment and the control in
passage 2 or 5 (pairwise Wilcoxon test: p = `r myround(wt[1, 3], 1)`).  In addition, there was no
correlation between richness and CH~4~ oxidation rate across the two treatments in passage 5 (Spearman's rho = `r myround(s5[1,1], 1)`, p = `r myround(s5[1,3], 1)`).

@@

This relationship between microbiome variation and ecosystem function variation
is analogous to the concept of "heritability" [@lynch1998] used by quantitative
geneticists, or more precisely the concept of "microbiability" [@difford2016] proposed
by microbiome scientists who study host-associated microbiomes. Although rarely
used in the study of environmental microbiomes, this concept could be very
useful for understanding and manipulating microbially-mediated functions in a
variety of ecosystems. In our experiment, variation in microbiome taxonomic
composition statistically explained (i.e., was associated with) 
`r myround(tidy(rh_model)[[2, 2]]*100, 1)`\% of the
variation we observed in the rate of CH~4~ oxidation.  This suggests that
there is substantial potential for altering this ecosystem function through
microbiome manipulation in this soil. It is very likely that the
“environmental microbiability” will be different for other ecosystem functions
in this soil and for CH~4~ oxidation in other soils. However, our experiment
demonstrates that this relationship is measurable and provides an example of how
this can be accomplished.

Given that variation in the microbiome is driving substantial variation in
CH~4~ oxidation rate in our system, we next wanted to determine which aspects of the
microbiome best explain this contribution of microbiome variation to
CH~4~ oxidation rate. There are three inter-related ways that microbiomes could have responded
to selection in this experiment: gain or loss of taxa, changes in the
relative abundances of taxa, or changes within the genomes of the constituent
taxa. We surveyed microbiome variation via 16s rRNA ribotyping in our
experiment, which allowed us to deeply sample taxonomic diversity but did not
allow us to directly address whether taxa in this experiment evolved genomic changes
as a result of selection. However, if such genomic changes resulted in increased
persistence or abundance of the population with these changes, this would be
detectable. Therefore, we will focus on the first
two possibilities.

@@

Even though we observed an increase in CH~4~ oxidation rate in the selection
treatment and a difference in composition between the two treatments, we did
not observe an increase in the relative abundance of methanotrophs. This was
surprising given that CH~4~ consumption is not a common trait among microbes
and that it is often assumed that the rate of an ecosystem function is limited
by the final enzymatic step in the underlying metabolic pathway [@rocca2015]. In certain
ecosystems, CH~4~ production and consumption are correlated with the
abundance of methanogens and methanotrophs as estimated from marker genes
[@meyer2020; @meyer2020a]. However, our results suggest that
ecosystem-scale CH~4~ oxidation rates can be altered by non-methanotrophs,
perhaps through ecological interactions with methanotrophic species. This
suggests that simple assumptions about how microbes contribute to rate variation
in ecosystem function may not apply universally, and it demonstrates the
importance of using biologically "agnostic" approaches (that make few starting
assumptions) to linking microbial taxa to ecosystem functions [@morris2020].
Artificial ecosystem selection is an important example of such an approach.

@@

# Supplement

@@

`r figures("herit", paste0("Divergence between the positive selection line and the control line over the cumulative selection differential. Values are CH~4~ oxidation rate as the additive inverse of the first-order exponential decay constant k (day^-1^). Regression of divergence on cumulative selection differential provides an estimate of realized heritability (h^2^ &pm; SE) and the slope of the regression was ", myround(tidy(rh_model)[[2, 2]], 2), " &pm; ", myround(tidy(rh_model)[[2, 3]], 2), ", though this was not significant (F~", glance(rh_model)$df,  ",", glance(rh_model)$df.residual,  "~ = ", myround(glance(rh_model)$statistic, 2),  ", p =  ", myround(glance(rh_model)$p.value, 2),  ")."))`

>>>>

Microbiomes mediate important ecosystem functions, yet it has proven difficult
to determine the relationship between microbiome composition and the rate of
ecosystem functions. This challenge remains because it is difficult to
manipulate microbiome composition directly, we often cannot know _a priori_ which
microbiome members influence the rate of an ecosystem function, and
microbiomes can covary strongly with other drivers of ecosystem function, such
as the environment. To address these challenges, we imposed artificial
selection on whole soil ecosystems over multiple generations to select for microbial communities
with a high rate of CH~4~ oxidation. This approach
is potentially powerful because it is biologically "agnostic" in 
that it makes few assumptions about which taxa are important to function, and
repeated passaging with fresh substrate weakens the covariance between microbes
and the environment. As a response to selection, we observed a
 `r myround(per_slope, 1)`\% increase in
CH~4~ oxidation rate per passage relative to a control that experienced
random selection. We
estimated that `r myround(tidy(rh_model)[[2, 2]]*100, 1)`\% of the variation in
CH~4~ oxidation
rate in these soils can be attributed to microbiome variation (though this was not significant). We also found
that selection did not enrich for known CH~4~ oxidizers; instead, `r length(da_fams)` families not
known to oxidize CH~4~, including *Fimbriimonadaceae*, *Cytophagaceae*, and
*Diplorickettsiaceae*, were enriched by selection. This result is in contrast to the typical assumption that
the rate of an ecosystem function is limited by the final step in the associated
microbial pathway. Our study demonstrates that variation in microbiome
composition can contribute to variation in the rate of ecosystem function 
independent of the environment and that this may not always be limited by the final step
in a pathway. This suggests
that manipulating microbiome composition directly without altering the environment 
could be a viable strategy for managing
ecosystem functions.

@@

In this study, we sought to build on the observations of comparative and
manipulative studies by applying a different approach to the question of whether
microbiome variation contributes to variation in the rate of an ecosystem
function. We used artificial ecosystem selection to select for microbiomes that
performed a greater rate of ecosystem function [@swenson2000; @blouin2015;
@panke-buisse2015]. We then tested whether variation in the microbiome
contributed to variation in the rate of ecosystem function and identified
microbiome attributes that might explain this relationship. There are several
potential advantages to this approach for documenting the direct contribution of
the microbiome to variation in ecosystem function and for investigating the
mechanisms underlying those relationships. Through repeated passaging of microbiomes
in a common environment, we can weaken the covariance between microbes and the
environment by repeatedly diluting the influence of variation in abiotic
conditions. In addition, our approach eliminates the need to generate microbiome
variation through methods that are confounded with biomass or cell size. Lastly,
by comparing our artificially selected community to a control community
resulting from random selection, we can both control for changes in the
environment over time and identify genes or taxa that are associated with the
ecosystem function under selection.

@@

We performed an artificial ecosystem selection experiment (*sensu* @swenson2000)
by passaging replicate soil microbiomes. The trait we selected on was soil CH~4~
oxidation rate.  Soil microcosms were incubated at room temperature in sealed
500 mL glass jars with a rubber septum for gas sampling. Each jar was sterilized
with 70\% ethanol and was composed of 45 g of autoclaved artificial potting mix,
5 g of living soil inoculum, and 3.5 mL of sterile deionized water to bring the
soil to 60\% of field capacity. The potting mix consisted of bark fines, peat
moss, pumice, sand, composted manure, and biochar (Lane Potting Mix, Lane Forest
Products, Eugene, OR). The initial soil microbiome inoculum was sampled from the
top 10 cm of an upland mineral soil under a deciduous forest ecosystem near the
University of Oregon campus in Eugene, OR, USA. Each jar
was capped and injected with 4.3 mL of 99\% CH~4~, which produced a mean
headspace concentration of 763.9 ppm (SD = 183.1). Twice per week, jars were
flushed in a biosafety cabinet (to avoid contamination) and respiked with CH~4~ 
to maintain aerobic conditions and elevated CH~4~ concentrations.

For the selection experiment, we created two lines of soil microcosms with 12
jars each: a control line with random selection and an experimental line with
directional selection for greater soil CH~4~ oxidation rate. The selection line
underwent positive selection where the two or three jars
with the highest CH~4~ oxidation rate were homogenized to inoculate the next set
of jars. The control line underwent random selection where an equal number
of jars as the selection line were chosen at random to inoculate the next set of
jars. The number of jars chosen was based on the distribution of fluxes among
the positive jars: three jars were chosen in all generations except for passage 3
where two jars were selected. The experiment was carried out over five passages
with an average
incubation time per generation of four weeks. 
Methane oxidation rates were determined at the end of
the incubation period and selection was performed. For each treatment, the
selected jars were homogenized and 5 g of the homogenized soil was used as the
living soil inoculum for the next generation. The next set
of jars were created in an identical manner to the first generation with fresh
autoclaved potting mix and the same moisture and CH~4~ content.

@@

Methane oxidation rates were determined after flushing and spiking jars to
1000 ppm CH~4~. Headspace samples of 1 mL were collected from
each jar immediately after spiking and then at time points 3, 6, 24, and 48
hours for a 5-point curve. Samples were immediately injected into a SRI model
8610C gas chromatograph equipped with a flame ionization detector (SRI
Instruments, Torrance, CA, USA) to determine the headspace CH~4~ concentration.
We applied a first-order exponential decay function to determine the rate
constant (k, units = d^−1^; i.e., dCH~4~/dt = k[CH~4~]) of the exponential
decrease in CH~4~. Oxidation rates are presented as the additive inverse of
$k$ (i.e., $-k$) so that a more positive value represents a greater
oxidation rate. The jars selected to inoculate passage three for the positive 
selection treatment had
the lowest CH~4~ oxidation rate of the twelve jars due to a calculation error
in the rate constant.

@@

Bioinformatics processing was performed in 'R' [@rcoreteam2018].  Demultiplexed
sequencing reads were denoised using 'DADA2' to generate a table of amplicon
sequence variants (ASVs) [@callahan2016]. Taxonomic assignment was performed
using the Ribosomal Database Project naive Bayesian classifier [@wang2007]. The
presence of contaminants was evaluated using both the prevalence and frequency
methods from 'DECONTAM' by comparing samples to extraction controls of 
water [@davis2018]. Decontam identified 
`r length(which(contamdf$contaminant))` potential contaminants based on
prevalence and frequency. Visual inspection of abundance-concentration plots
indicated that `r length(contaminants)` of these were likely contaminants and
these ASVs were removed. Amplicon sequence variants that were assigned
chloroplast or mitochondria taxonomy were removed prior to analysis.

@@

Statistical analyses were performed in 'R' [@rcoreteam2018]. To determine
whether there was a significant change in CH~4~ oxidation rate as a response to
selection, we tested a difference in slopes between the selection and control
lines. Residuals did not meet the assumptions of constant variance
and normal distribution. Therefore, CH~4~ oxidation rates were log~10~
transformed prior to analysis. Following transformation, these assumptions
were met. First, we tested if there was a difference of
slopes between the selection line and the control based on the interaction
between passage and treatment. To test the interaction, we fit two nested models
with and without the interaction term and compared them using an F-test with the
'anova' function. We then present the slopes for each treatment, which
represented the change in CH~4~ oxidation rate per passage as a response to
selection.

We estimated the proportion of variation in CH~4~ oxidation rate due to
variation in the microbiome as the regression of divergence between the positive
line and the control on the cumulative selection differential [@falconer1996].
This estimate is analogous
to estimates of "microbiability" from the animal breeding literature, which
quantifies the
variation in a host trait that is due to microbiome variation [@he2022]. 
The slope of the regression of divergence on cumulative selection differential
provides an estimate of realized microbiability (h^2^ &pm; SE).
Divergence was calculated as the mean CH~4~ oxidation rate of the positive
treatment minus the mean CH~4~ oxidation rate of the control in each
passage. The selection differential was calculated as the difference between
the mean of the three selected jars and the mean of all twelve jars in a passage.
Cumulative selection differential was calculated as the sum of the selection
differential from all preceding selection events. We then regressed cumulative
divergence on cumulative selection differential using the 'lm' function. We report
the slope as percent change by back-calculating the percent change from the
log-transformed data into the original units using the formula $(10^\beta - 1) * 100$
where $\beta$ is the slope.

@@

To identify taxa that responded to selection on CH~4~ oxidation rate, we tested
differential abundance between the two treatments in passage 5. We first
grouped ASVs at the family level. We chose this level of agglomeration
because CH~4~ oxidation
is a relatively deeply conserved function [@martiny2013] and is restricted to
a handful of bacterial and archaeal families [@kang2019]. Therefore, we are most
likely to detect an enrichment of methanotrophs at this taxonomic scale.
Any ASVs that lacked a family-level
taxonomic assignment were grouped at a higher taxonomic level. We then subset the
samples in Passage 5 and removed all families with a prevalence of less than
10\% in either treatment. We used three methods for testing differential
abundance: ANCOM-II, ALDEx2, and CORNCOB 
[@kaul2017; @lin2020; @fernandes2014; @martin2020]. We then identified the
consensus taxa that were significant with all three tests and plotted their
relative abundances. For ANCOM-II, we used the 'ancom' function in the
'ANCOM-BC' package with a cutoff of W = 0.7 [@kaul2017; @lin2020]. For ALDEx2,
we used the 'aldex' function in the 'ALDEx2' package with Welch's t-test and we
used an effect size of 1 as our significance threshold [@fernandes2014].
Finally, we used CORNCOB with the 'differentialTest' function in the 'corncob'
package with the Wald test and without bootstrapping [@martin2020]. Lastly, to
test differentially abundant methanotrophs, we subset all ASVs within
methanotrophic families and tested their differential abundance aggregated at
the family and genus level using 'corncob'. For each test, p-values were adjusted
for multiple testing by controlling the false discovery rate using the 
Banjamini-Hochberg procedure [@benjamini1995].

@@

To estimate the proportion of variation in CH~4~ oxidation rate due to
variation in microbiome composition--i.e., microbiability (as described in the Methods; [@he2022]--we regressed divergence between the
positive selection treatment and the control against the cumulative selection
differential. The microbiability was `r myround(tidy(rh_model)[[2, 2]], 2)` &pm;
`r myround(tidy(rh_model)[[2, 3]], 2)`, though this was not significant (F~`r glance(rh_model)$df`,`r glance(rh_model)$df.residual`~ = 
`r myround(glance(rh_model)$statistic, 2)`, 
p = `r myround(glance(rh_model)$p.value, 2)`).

@@

Median ASV richness decreased from `r myround(rare_median[1, 2], 1)` (`r myround(rare_median[1, 3], 1)`) in passage 2 to `r myround(rare_median[2, 2], 1)` (`r myround(rare_median[2, 3], 1)`) in passage 5 (Kruskal-Wallace test: $\chi^2$ = 
`r myround(kt[1,1], 1)`, df = `r kt[1,3]`, p < 0.001; pairwise Wilcoxon test: p < 0.001). However, there was no
difference in richness between the selection treatment and the control in
passage 2 or 5 (pairwise Wilcoxon test; Passage 2: p = `r myround(wt[1, 3], 2)`,
Passage 5: p = `r myround(wt[6, 3], 2)`).  In addition, there was no
correlation between richness and CH~4~ oxidation rate across the two treatments in passage 5 (Spearman's rho = `r myround(s5[1,1], 1)`, p = `r myround(s5[1,3], 1)`).

@@

The relationship between microbiome variation and ecosystem function variation
is analogous to the concept of "heritability" [@lynch1998] used by quantitative
geneticists, or more precisely the concept of "microbiability" [@he2022] proposed
by microbiome scientists who study host-associated microbiomes. Although rarely
used in the study of environmental microbiomes, this concept could be very
useful for understanding and manipulating microbially-mediated functions in a
variety of ecosystems. In our experiment, variation in microbiome taxonomic
composition statistically explained (i.e., was associated with) 
`r myround(tidy(rh_model)[[2, 2]]*100, 1)`\% of the
variation we observed in the rate of CH~4~ oxidation, though this was not
significant. However, we did observe a significant divergence between the 
positive selection and control lines, which suggests that the imposed
selection and passaging of microbiomes was sufficient to generate
variation in soil CH~4~ oxidation rate. Future studies with
greater replication could more precisely estimate the microbiability.
This suggests that
there is substantial potential for altering this ecosystem function through
microbiome manipulation in this soil. It is very likely that the
“environmental microbiability” will be different for other ecosystem functions
in this soil and for CH~4~ oxidation in other soils. However, our experiment
demonstrates that this relationship is measurable and provides an example of how
this can be accomplished.

We next wanted to determine which aspects of the
microbiome might explain the divergence in CH~4~ oxidation rate between the two treatments.
There are three inter-related ways that microbiomes could have responded
to selection in this experiment: gain or loss of taxa, changes in the
relative abundances of taxa, or changes within the genomes of the constituent
taxa. We surveyed microbiome variation via 16s rRNA ribotyping in our
experiment, which allowed us to deeply sample taxonomic diversity but did not
allow us to directly address whether taxa in this experiment evolved genomic changes
as a result of selection. However, if such genomic changes resulted in increased
persistence or abundance of the population with these changes, this would be
detectable. Therefore, we will focus on the first
two possibilities.

@@

Even though we observed an increase in CH~4~ oxidation rate in the selection
treatment and a difference in composition between the two treatments, we did
not observe an increase in the relative abundance of known methanotrophs. This was
surprising given that CH~4~ consumption is not a common trait among microbes
and that it is often assumed that the rate of an ecosystem function is limited
by the final enzymatic step in the underlying metabolic pathway [@rocca2015]. In certain
ecosystems, CH~4~ production and consumption are correlated with the
abundance of methanogens and methanotrophs as estimated from marker genes
[@meyer2020; @meyer2020a]. However, our results suggest that in this system
ecosystem-scale CH~4~ oxidation rates can be altered by non-methanotrophs,
perhaps through ecological interactions with methanotrophic species, or by
unknown methanotrophs. This
suggests that simple assumptions about how microbes contribute to rate variation
in ecosystem function may not apply universally, and it demonstrates the
importance of using biologically "agnostic" approaches (that make few starting
assumptions) to linking microbial taxa to ecosystem functions [@morris2020].
Artificial ecosystem selection is an important example of such an approach.

@@

@@
